---
title: "Random Forest - biological implications of aging"
author: "Chelsea Weibel"
date: "4/17/2024"
output:
  html_document:
    theme: sandstone
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages
```{r}
library(tidyverse)
library(gridExtra)
library(lme4)
library(lmerTest)
library(survival)
require(RColorBrewer)
library(survminer)
```


# Loading data
```{r}
# rf data
rf_avg_predictions <- read_csv("../1_data/rf_avg_predictions.csv")

adversity_covariates <- read_csv("../1_data/adversity_covariates.csv")

# combining data frames and selecting columns of interest
data_bioage <- rf_avg_predictions %>% 
  select(anon_sname, age_year, bio_age = avg_bio_age) %>% 
  left_join(select(adversity_covariates, anon_sname, 
                   birth, age_year, 
                   statdate, status, age_at_statdate,
                   mom_rank_birth, group_size_birth, mom_sci,
                   mom_rank_birth_cat, group_size_birth_cat, mom_sci_cat,
                   mom_death_cat, competing_sibling_cat, drought_cat,
                   cumulative_adversity, cumulative_adversity_three_plus,
                   avg_ordrank, avg_proprank,
                   rain_anom,
                   avg_group_size),
            by=c("anon_sname", "age_year"))
```


# Data formatting

## Calculating delta age
```{r}
###################### DELTA AGE
# creating the delta age columns
# biological - chronological (pos = old, neg = young)
data_bioage <- data_bioage %>% 
# calculating delta age
  mutate(delta_age = bio_age - age_year)


# Average delta age per individual
ind_data <- data_bioage %>% 
  group_by(anon_sname, birth, statdate, status, age_at_statdate) %>% 
  dplyr::summarize(avg_delta_age = as.numeric(mean(delta_age)),
                   avg_age = as.numeric(mean(age_year)))
```

## Combining with EA and other covariates
```{r}
# creating an ea file for each individual (removes duplicate rows for each individual)
ea <- adversity_covariates %>% 
  select(anon_sname, mom_rank_birth:cumulative_adversity_three_plus) %>% 
  distinct() %>% 
  mutate(cumulative_adversity_three_plus_categorical = as.factor(cumulative_adversity_three_plus))

covariates_avg <- adversity_covariates %>% 
  select(anon_sname, avg_ordrank, avg_proprank, rain_anom, avg_group_size) %>% 
  group_by(anon_sname) %>% 
  dplyr::summarize(overall_avg_ordrank = mean(avg_ordrank),
            overall_avg_proprank = mean(avg_proprank),
            overall_rain_anom = mean(rain_anom),
            overall_avg_group_size = mean(avg_group_size))

ind_data <- ind_data %>% 
  left_join(select(ea, anon_sname, mom_rank_birth, group_size_birth, mom_sci,
                   mom_rank_birth_cat, group_size_birth_cat, mom_sci_cat,
                   mom_death_cat, competing_sibling_cat, drought_cat,
                   cumulative_adversity:cumulative_adversity_three_plus_categorical),
            by=c("anon_sname")) %>% 
  left_join(covariates_avg, by="anon_sname")

# creating a subset data frame where there's no missing CEA info
ind_data_cea <- ind_data %>% 
  filter(!is.na(cumulative_adversity_three_plus))
# 256 individuals

ind_data_cea %>% 
  group_by(cumulative_adversity) %>% 
  dplyr::summarize(n = n()) %>% 
  mutate(per = n / 256)
```


# Does adversity predict delta age?

## Average per individual

### Cumulative adversity
```{r}
# model
cea_mod <- lm(avg_delta_age ~ cumulative_adversity_three_plus + 
                avg_age + overall_avg_proprank + overall_rain_anom + overall_avg_group_size, data=ind_data_cea)
summary(cea_mod)
```


#### Plot
```{r}
# CEA plots with residuals as y-axis

# redoing model without CEA in it
no_cea_mod <- lm(avg_delta_age ~ avg_age + overall_avg_proprank + overall_rain_anom + overall_avg_group_size, data=ind_data_cea)
summary(no_cea_mod)

# extracting the residuals
ind_data_cea$delta_age_resid_no_cea = resid(no_cea_mod)

# scatter
ind_data_cea %>% 
  ggplot(aes(x=cumulative_adversity_three_plus, y=delta_age_resid_no_cea)) +
  geom_jitter() +
  geom_smooth(method="lm") +
  theme_bw() +
  xlab("cumulative adversity") +
  ylab("delta age residuals") +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))

# boxplot
ind_data_cea %>% 
  ggplot(aes(x=cumulative_adversity_three_plus_categorical, y=delta_age_resid_no_cea)) +
  geom_jitter(aes(color=cumulative_adversity_three_plus_categorical)) +
  geom_boxplot(outlier.shape=NA, fill=NA) +
  theme_bw() +
  xlab("Cumulative adversity") +
  ylab("Delta age, controlling for chronological age") +
  theme(legend.position="none") +
  scale_color_manual(values=c("0" = "#3E76A7",
                              "1" = "#A5C6D9", 
                              "2" = "#DE9B34",
                              "3" = "#B32A06")) +
  scale_x_discrete(labels=c("0" = "0", "1" = "1",
                              "2" = "2", "3" = "3+"))
```

### Individual sources of adversity
```{r}
# model
iea_mod <- lm(avg_delta_age ~ mom_rank_birth + group_size_birth + mom_sci + 
                mom_death_cat + competing_sibling_cat + drought_cat + 
                avg_age +
                overall_avg_proprank + overall_rain_anom + overall_avg_group_size, data=ind_data_cea)
summary(iea_mod)
```

#### All categorical
```{r}
# model
iea_mod_allcat <- lm(avg_delta_age ~ mom_rank_birth_cat + group_size_birth_cat + mom_sci_cat + 
                mom_death_cat + competing_sibling_cat + drought_cat + 
                avg_age +
                overall_avg_proprank + overall_rain_anom + overall_avg_group_size, data=ind_data_cea)
summary(iea_mod_allcat)
```


#### Residuals adversity
```{r}
# plots with residuals for y axis

# first redoing the model without adversity so I can take residuals
no_iea_mod <- lm(avg_delta_age ~ avg_age +
                overall_avg_proprank + overall_rain_anom + overall_avg_group_size, data=ind_data_cea)
summary(no_iea_mod)

# extracting the residuals
ind_data_cea$delta_age_resid_no_iea = resid(no_iea_mod)
```


## Yearly measure

```{r}
data_bioage_cea <- data_bioage %>% 
  filter(!is.na(cumulative_adversity_three_plus))
```

### Cumulative adversity (Table S6)
```{r}
# model
cea_mod_yearly <- lmer(delta_age ~ cumulative_adversity_three_plus + 
                age_year + avg_proprank + rain_anom + avg_group_size + 
                  (1|anon_sname),
                data=data_bioage_cea)
summary(cea_mod_yearly)
```

#### Figure 4
```{r, fig.height=4, fig.width=6}
cea_mod_yearly_nocea <- lmer(delta_age ~ age_year + avg_proprank + rain_anom + avg_group_size +
                               (1|anon_sname),
                             data=data_bioage_cea)
data_bioage_cea$cea_yearly_resid = resid(cea_mod_yearly_nocea)


cea_delta_age_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=as.factor(cumulative_adversity_three_plus), y=cea_yearly_resid)) +
  geom_jitter(aes(color=as.factor(cumulative_adversity_three_plus))) +
  geom_boxplot(outlier.shape=NA, fill=NA) +
  theme_bw() +
  xlab("Cumulative adversity") +
  ylab("Delta age, controlling for covariates") +
  theme(legend.position="none") +
  scale_color_manual(values=c("0" = "#3E76A7",
                              "1" = "#A5C6D9", 
                              "2" = "#DE9B34",
                              "3" = "#B32A06")) +
  scale_x_discrete(labels=c("0" = "0", "1" = "1",
                              "2" = "2", "3" = "3+"))
cea_delta_age_yearly
```

### Individual sources of adversity (Table S7)
```{r}
# model
iea_mod_yearly <- lmer(delta_age ~ mom_rank_birth + group_size_birth + mom_sci + 
                mom_death_cat + competing_sibling_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), data=data_bioage_cea)
summary(iea_mod_yearly)
```

#### Figure S10
```{r}
# plots with residuals for y axis, creating a separate model for each source of adversity
# MOM RANK
no_mom_rank_mod_yearly <- lmer(delta_age ~ group_size_birth + mom_sci + mom_death_cat + competing_sibling_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), data=data_bioage_cea)
summary(no_mom_rank_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_mom_rank_yearly = resid(no_mom_rank_mod_yearly)

mom_rank_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=mom_rank_birth, y=delta_age_resid_no_mom_rank_yearly)) +
  geom_jitter() +
  geom_smooth(method="lm") +
  theme_bw() +
  xlab("Maternal rank") +
  ylab("Residuals of delta age") +
  ggtitle("Maternal rank") +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))


# GROUP SIZE
no_goup_size_mod_yearly <- lmer(delta_age ~ mom_rank_birth + mom_sci + mom_death_cat + competing_sibling_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), 
                data=data_bioage_cea)
summary(no_goup_size_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_group_size_yearly = resid(no_goup_size_mod_yearly)

group_size_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=group_size_birth, y=delta_age_resid_no_group_size_yearly)) +
  geom_jitter() +
  geom_smooth(method="lm") +
  theme_bw() +
  xlab("Group size") +
  ylab("Residuals of delta age") +
  ggtitle("Group size") +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))


# SCI
no_sci_mod_yearly <- lmer(delta_age ~ mom_rank_birth + group_size_birth + mom_death_cat + competing_sibling_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), 
                data=data_bioage_cea)
summary(no_sci_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_sci_yearly = resid(no_sci_mod_yearly)

maternal_sci_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=mom_sci, y=delta_age_resid_no_sci_yearly)) +
  geom_jitter() +
  geom_smooth(method="lm") +
  theme_bw() +
  xlab("Maternal social isolation") +
  ylab("Residuals of delta age") +
  ggtitle("Maternal social isolation") +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))


# MATERNAL DEATH
no_maternal_death_mod_yearly <- lmer(delta_age ~ mom_rank_birth + group_size_birth + mom_sci + competing_sibling_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname),
                data=data_bioage_cea)
summary(no_maternal_death_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_maternal_death_yearly = resid(no_maternal_death_mod_yearly)

maternal_death_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=as.factor(mom_death_cat), y=delta_age_resid_no_maternal_death_yearly)) +
  geom_jitter(aes(color=as.factor(mom_death_cat))) +
  geom_boxplot(outlier.shape=NA, fill=NA) +
  theme_bw() +
  xlab("Maternal death") +
  ylab("Residuals of delta age") +
  ggtitle("Maternal death") +
  scale_color_manual(values=c("0" = "#3E76A7",
                              "1" = "#B32A06")) +
  theme(legend.position="none") +
  scale_x_discrete(labels = c("0" = "no",
                              "1" = "yes")) +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))


# COMPETING SIBLING
no_competing_sibling_mod_yearly <- lmer(delta_age ~ mom_rank_birth + group_size_birth + mom_sci + mom_death_cat + drought_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), 
                data=data_bioage_cea)
summary(no_competing_sibling_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_competing_sibling_yearly = resid(no_competing_sibling_mod_yearly)

competing_sibling_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=as.factor(competing_sibling_cat), y=delta_age_resid_no_competing_sibling_yearly)) +
  geom_jitter(aes(color=as.factor(competing_sibling_cat))) +
  geom_boxplot(outlier.shape=NA, fill=NA) +
  theme_bw() +
  xlab("Close-in-age sibling") +
  ylab("Residuals of delta age") +
  ggtitle("Close-in-age sibling") +
  scale_color_manual(values=c("0" = "#3E76A7",
                              "1" = "#B32A06")) +
  theme(legend.position="none") +
  scale_x_discrete(labels = c("0" = "no",
                              "1" = "yes")) +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))


# DROUGHT
no_drought_mod_yearly <- lmer(delta_age ~ mom_rank_birth + group_size_birth + mom_sci + mom_death_cat + competing_sibling_cat + 
                age_year + avg_proprank + rain_anom + avg_group_size +
                  (1|anon_sname), 
                data=data_bioage_cea)
summary(no_drought_mod_yearly)

# extracting the residuals
data_bioage_cea$delta_age_resid_no_drought_yearly = resid(no_drought_mod_yearly)

rainfall_rf_resid2_yearly <- data_bioage_cea %>% 
  ggplot(aes(x=as.factor(drought_cat), y=delta_age_resid_no_drought_yearly)) +
  geom_jitter(aes(color=as.factor(drought_cat))) +
  geom_boxplot(outlier.shape=NA, fill=NA) +
  theme_bw() +
  xlab("Drought") +
  ylab("Residuals of delta age") +
  ggtitle("Drought") +
  scale_color_manual(values=c("0" = "#3E76A7",
                              "1" = "#B32A06")) +
  theme(legend.position="none") +
  scale_x_discrete(labels = c("0" = "no",
                              "1" = "yes")) +
  scale_y_continuous(limits=c(-2,3), breaks=seq(-2, 3, 1))
```

```{r, fig.width=10, fig.height=6}
ind_adv_delta_age_yearly <- grid.arrange(mom_rank_rf_resid2_yearly, group_size_rf_resid2_yearly, maternal_sci_rf_resid2_yearly, maternal_death_rf_resid2_yearly, competing_sibling_rf_resid2_yearly, rainfall_rf_resid2_yearly, nrow=2)
```


# Survival analysis
```{r}
survival <- ind_data %>%
# creating a censored/dead column
# 2 = dead, 1 = censored
  mutate(censor_dead = ifelse(status == 1, 2, 1))
# 162 individuals dead, 157 censored

```

## Delta age

### Average across lifespan
```{r}
survival %>% 
  ggplot(aes(x=avg_age, y=avg_delta_age)) +
  geom_point()


# running a linear model of avg_delta_age and avg age so that we can control for avg age without having it in the model
delta_age_resid_mod <- lm(delta_age ~ age_year, data = data_bioage)
# extracting the residuals
data_bioage$delta_age_resid = resid(delta_age_resid_mod)

# taking the average across individuals
avg_df <- data_bioage %>% 
  group_by(anon_sname) %>% 
  dplyr::summarize(avg_delta_age_resid = mean(delta_age_resid))


# and then merging with the survivals df
survival <- survival %>% 
  left_join(avg_df)


# running the model
delta_age_survival_model <- coxph(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid, data=survival)
print(summary(delta_age_survival_model),digits=10)

# testing proportional hazards assumption
cox.zph(delta_age_survival_model)
# assumptions are met




# creating quantiles for delta age residuals
survival <- survival %>% 
  mutate(avg_delta_age_resid_quant = case_when(
    avg_delta_age_resid <= unname(quantile(survival$avg_delta_age_resid, 0.25)) ~ 1,
    avg_delta_age_resid > unname(quantile(survival$avg_delta_age_resid, 0.25)) & 
                      avg_delta_age_resid <= unname(quantile(survival$avg_delta_age_resid, 0.50)) ~ 2,
    avg_delta_age_resid > unname(quantile(survival$avg_delta_age_resid, 0.50)) & 
                      avg_delta_age_resid <= unname(quantile(survival$avg_delta_age_resid, 0.75)) ~ 3,
    avg_delta_age_resid > unname(quantile(survival$avg_delta_age_resid, 0.75)) ~ 4))

# quartile cutoffs
quantile(survival$avg_delta_age_resid)


# running the model
delta_age_survival_model_quant <- coxph(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid_quant, data=survival)
print(summary(delta_age_survival_model_quant),digits=10)
cox.zph(delta_age_survival_model_quant)
# assumptions are met

survfit(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid_quant, data=survival)


# adversity colors
require(RColorBrewer)
cols<-rev(brewer.pal(4,"RdYlBu"))

# quartile colors
cols_quart <- c("#456990", "#88AB75", "#F39237", "#D81159") 

# creating and saving this survival plot
#pdf("plots/FigS1.pdf", height = 6, width = 5)
plot(survfit(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid_quant, data=survival),
     mark.time=F,
     col=cols_quart,
     frame.plot=F,
     lwd=3,
     xlab='Age (years)',
     ylab='Proportion surviving',
     cex.lab=1.5,
     cex.axis=1.5,
     #firstx=4,
     xlim=c(4,30))
legend("topright",c("Q1 - look youngest for age","Q2 - look slightly young for age","Q3 - look slightly old for age","Q4 - look oldest for age"),fill=cols_quart,bty="n",cex=0.9,border=NA,title='Lifetime delta age quartile')
```


#### Figure 3A

```{r, fig.height=4, fig.width=5.5}
lifetime_fit <- survfit(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid_quant, data=survival)
lifetime_plot <- ggsurvplot(lifetime_fit, data = survival,
           ggtheme = theme_bw(base_size = 14),
          # surv.median.line = "hv",
           legend.title = "Delta age quartile",
           legend.labs = c("Q1", "Q2", "Q3", "Q4"),
           palette = c("#456990", "#88AB75", "#F39237", "#D81159"),
           xlab = "Age (years)",
          ## Add risk table
 #risk.table = TRUE,
 #tables.height = 0.2,
 #tables.theme = theme_cleantable(),
          #conf.int = T,
          #pval = T,
          legend = "right",
          censor = F,
 xlim = c(4,28)) +
  ggtitle("Lifetime delta age")
lifetime_plot
```


#### Average across lifespan, including EA (Table S8)
```{r}
# first subsetting the data
survival_ea <- survival %>% 
  filter(!is.na(cumulative_adversity_three_plus)) %>% 
  select(-avg_delta_age_resid, -avg_delta_age_resid_quant)

# came as data_bioage_cea, but I already made it, so we'll use it
data_bioage_ea <- data_bioage %>% 
  filter(anon_sname %in% survival_ea$anon_sname) %>% 
  select(-delta_age_resid)


# running a linear model of delta_age and age so that we can control for age without having it in the model
delta_age_resid_mod2 <- lm(delta_age ~ age_year, data = data_bioage_ea)
# extracting the residuals
data_bioage_ea$delta_age_resid_ea = resid(delta_age_resid_mod2)


# taking the average across individuals
avg_df_ea <- data_bioage_ea %>% 
  group_by(anon_sname) %>% 
  dplyr::summarize(avg_delta_age_resid_ea = mean(delta_age_resid_ea))


# and then merging with the survivals df
survival_ea <- survival_ea %>% 
  left_join(avg_df_ea)


# checking for correlations and VIF
cor(survival_ea$avg_delta_age_resid_ea, survival_ea$cumulative_adversity_three_plus)
survival_ea %>% 
  ggplot(aes(x=cumulative_adversity_three_plus, y=avg_delta_age_resid_ea)) +
  geom_jitter() +
  geom_smooth(method="lm")



# running the model
delta_age_survival_model2 <- coxph(Surv(age_at_statdate, censor_dead) ~ avg_delta_age_resid_ea + cumulative_adversity_three_plus, data=survival_ea)
print(summary(delta_age_survival_model2),digits=10)
cox.zph(delta_age_survival_model2)
```


### Time-varying
https://stats.oarc.ucla.edu/r/dae/mixed-effects-cox-regression/
```{r}
data_bioage <- data_bioage %>% 
# creating an end date for each row
  mutate(start = age_year,
         end = age_year + 1) %>% 
# creating a survival column
# IDing last rows
  mutate(last_row = age_at_statdate - end) %>% 
  mutate(survival = ifelse(last_row < 0 & status == 1, 1, 0)) %>% 
# we have info during the last year of life for 113 individuals
# since we don't have data for the last year of life for all individuals (partly because they didn't survive the whole year so we don't have enough info for them, going to count the last year of life as any year when it's 1/2 year before death or later)
# first calculating the min last_row for each individual
  group_by(anon_sname) %>% 
  mutate(min_last_row = min(last_row)) %>% 
  mutate(survival2 = ifelse(min_last_row == last_row & last_row < 0.5 & status == 1, 1, 0)) %>% 
# WE DECIDED WE WANT THE LAST ROW TO END WITH THE DATE THAT THEY DIED
  mutate(end2 = ifelse(survival2 == 1, age_at_statdate, end))
```


```{r}
# with the second survival column
survmod2 <- coxph(Surv(start, end2, survival2) ~ delta_age + age_year + cluster(anon_sname), data=data_bioage)
summary(survmod2)
```

#### With residuals (second survival column)
```{r}
survmod_resid <- coxph(Surv(start, end2, survival2) ~ delta_age_resid + cluster(anon_sname), data=data_bioage)
summary(survmod_resid)
cox.zph(survmod_resid)

summary(data_bioage$delta_age_resid)
```


#### Include EA (Table S9)
```{r}
data_bioage_ea <- data_bioage_ea %>% 
  left_join(select(data_bioage, anon_sname, age_year, start, end2, survival2), by = c("anon_sname", "age_year"))

survmod_resid_ea <- coxph(Surv(start, end2, survival2) ~ delta_age_resid_ea + cumulative_adversity_three_plus + cluster(anon_sname), data=data_bioage_ea)
summary(survmod_resid_ea)
cox.zph(survmod_resid_ea)
```


#### Categorical (second survival column)
```{r}
# CREATING A QUANTILES COLUMN AND MODEL SO I CAN PLOT THE EFFECTS
data_bioage <- data_bioage %>% 
  mutate(delta_age_resid_quant = case_when(
    delta_age_resid <= unname(quantile(data_bioage$delta_age_resid, 0.25)) ~ 1,
    delta_age_resid > unname(quantile(data_bioage$delta_age_resid, 0.25)) & 
                      delta_age_resid <= unname(quantile(data_bioage$delta_age_resid, 0.50)) ~ 2,
    delta_age_resid > unname(quantile(data_bioage$delta_age_resid, 0.50)) & 
                      delta_age_resid <= unname(quantile(data_bioage$delta_age_resid, 0.75)) ~ 3,
    delta_age_resid > unname(quantile(data_bioage$delta_age_resid, 0.75)) ~ 4))


data_bioage %>% 
  group_by(delta_age_resid_quant) %>% 
  dplyr::summarize(avg_delta_age_resid = mean(delta_age_resid),
            min_delta_age_resid = min(delta_age_resid),
            max_delta_age_resid = max(delta_age_resid))


# creating a categorical model so I can plot
survmod_resid_cat <- coxph(Surv(start, end2, survival2) ~ as.factor(delta_age_resid_quant) + cluster(anon_sname), data=data_bioage)
summary(survmod_resid_cat)
cox.zph(survmod_resid_cat)

survfit(Surv(start, end2, survival2) ~ as.factor(delta_age_resid_quant), data=data_bioage)

```

#### Figure 3B
```{r, fig.height=4, fig.width=5.5}
# https://rpkgs.datanovia.com/survminer/reference/ggsurvplot.html#plot-title-and-axis-labels
yearly_fit <- survfit(Surv(start, end2, survival2) ~ as.factor(delta_age_resid_quant), data=data_bioage)
yearly_plot <- ggsurvplot(yearly_fit, data = data_bioage,
           ggtheme = theme_bw(base_size = 14),
           #surv.median.line = "hv",
           legend.title = "Delta age quartile",
           legend.labs = c("Q1", "Q2", "Q3", "Q4"),
           palette = c("#456990", "#88AB75", "#F39237", "#D81159"),
           xlab = "Age (years)",
          ## Add risk table
 #risk.table = TRUE,
 #tables.height = 0.2,
 #tables.theme = theme_cleantable(),
          #conf.int = T,
          #pval = T,
          legend = "right",
          censor = F,
 xlim = c(4,28)) +
  ggtitle("Yearly delta age")
yearly_plot
```



