---
title: "Investigating NAs"
author: "Chelsea Weibel"
date: "4/17/2024"
output:
  html_document:
    theme: sandstone
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages
```{r}
library(tidyverse)
library(naniar)
library(visdat)
```

# Loading data
```{r}
traits <- read_csv("../1_data/traits.csv")
```

# Evaluating NAs

## Number of NAs per row
```{r}
traits$na_count <- apply(is.na(traits), 1, sum)


traits %>% 
  ggplot() +
  geom_histogram(aes(x=na_count), binwidth = 1) +
  theme_bw()
# going to use 34 as a cutoff point
```

## Removing rows with lots of NAs
```{r}
# for now, we want to keep individuals' last years of data
# but we end up removing these later on

traits <- traits %>% 
  mutate(yol = ifelse(end_date >= statdate, "final", "not_final")) %>% 
# filtering for last yol or rows with < 34 nas
  filter(yol == "final" | na_count < 34)
# now at 2470 (removed 807 rows)
```


## Figure S1 Inset
```{r}
nas_plot <- traits %>% 
  filter(na_count < 34) %>% # excluding final yol because we drop them later
  ggplot() +
  geom_histogram(aes(x=na_count), binwidth = 1) +
  theme_bw(base_size = 16) +
  ylab("Number of female-years of life") +
  xlab("Number of traits missing")

nas_plot
```

## How many cells were missing? (before backfill)
```{r}
missing_cells_before_backfill <- traits %>% 
  filter(na_count < 34)

n_missing_cells_before_backfill <- sum(is.na(missing_cells_before_backfill))
total_cells <- 2402 * 49

n_missing_cells_before_backfill/total_cells
```

## Number of NAs per trait
```{r}
# used for Table S4

missing_per_col <- missing_cells_before_backfill %>%
  ungroup() %>% 
  select(per_groom_resid, ni_per_forage_resid, ni_per_groom_resid, per_no_neighbors_resid, ni_per_no_neighbors_resid, ni_per_other_soc_resid, ni_per_rest_resid, 
         avg_resid_log_e, avg_resid_log_gc, avg_resid_log_p, avg_resid_log_th,
         limp_cat, major_health_event_cat, puncture_cat,
         avg_richness_resid, avg_log_trichuris_resid,
         avg_log_cycling_resid, any_off_survive, avg_preg_resid, any_misc,
         per_kc_away_none_resid, per_be_groomed_infant_resid, per_ks_suckle_resid, per_kc_support_resid, per_kc_dorsal_resid, per_kc_ventral_resid,
         eigen_resid, not_bonded_f_resid, not_bonded_m_resid, strongly_bonded_f_resid, total_grooming_partners_f_resid, total_grooming_partners_resid, weakly_bonded_f_resid, sci_f_resid, sci_f_dir_resid, sci_m_dir_resid, sci_f_rec_resid, sci_m_rec_resid,
         dsi_f_resid, recip_f_resid, recip_m_resid,
         agi_f_resid, agi_m_resid, agi_f_dir_resid, agi_f_rec_resid, agi_m_rec_resid,
         cat_abs_active_rank_change, rank_rel_any_daughter, avg_mom_rel_rank)  %>% 
  summarise_all(funs(sum(is.na(.)))) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(per = round((V1/2402)*100, 2)) %>% 
  mutate(lab = paste0(V1, " (", per, "%)"))
```


# Forward- and backward-filling

## Filling backward/forward for numeric factors (41)
```{r}
# creating lists of numeric traits
numeric_biomarkers <- traits %>% 
  select_if(is.numeric) %>% 
  select(contains("resid")) %>% 
  colnames()

# creating an output df for the new values
numeric_output <-data.frame(matrix(ncol = length(numeric_biomarkers), nrow = nrow(traits)))
colnames(numeric_output) <- numeric_biomarkers

# for each numeric trait, if the trait is NA, fill in with the next or previous or take the mean
# new output is in the numeric_output df
for (i in 1:length(numeric_biomarkers)) {
df <- traits %>% 
  group_by(anon_sname) %>% 
  select(anon_sname, as.name(numeric_biomarkers[i]))
colnames(df) <- c("anon_sname", "trait")

df <- df %>%
  mutate(next_trait = lead(trait),
         prev_trait = lag(trait)) %>% 
  mutate(new_trait = case_when(!is.na(trait) ~ trait,
                            is.na(trait) & is.na(prev_trait) & !is.na(next_trait) ~ next_trait,
                            is.na(trait) & !is.na(prev_trait) & is.na(next_trait) ~ prev_trait,
                            is.na(trait) & !is.na(prev_trait) & !is.na(next_trait) ~ 
                                            (prev_trait + next_trait)/2))

numeric_output[,i] <- df$new_trait
}
```

## Filling backward/forward for factors
```{r}
factor_biomarkers <- traits %>% 
  select(!contains("resid")) %>% 
  select(any_misc:puncture_cat) %>% 
  colnames()

# assigning a random a/b to each row so we can randomize whether we fill forwards or backwards for the factors
traits <- traits %>% 
  ungroup() %>% 
  mutate(group = sample(c('A','B'), size = nrow(traits), prob = c(0.5,0.5), replace = TRUE))
 
# creating an output df for the new values
factor_output <-data.frame(matrix(ncol = length(factor_biomarkers), nrow = nrow(traits)))
colnames(factor_output) <- factor_biomarkers

# for each factor, if the trait is NA, fill in with the next or previous
# new output is in the factor_output df
for (i in 1:length(factor_biomarkers)) {
df <- traits %>% 
  group_by(anon_sname) %>% 
  select(anon_sname, group, as.name(factor_biomarkers[i]))
colnames(df) <- c("anon_sname", "group", "trait")

df <- df %>%
  mutate(next_trait = lead(trait),
         prev_trait = lag(trait)) %>% 
  mutate(new_trait = case_when(!is.na(trait) ~ trait,
                            is.na(trait) & is.na(prev_trait) & !is.na(next_trait) ~ next_trait,
                            is.na(trait) & !is.na(prev_trait) & is.na(next_trait) ~ prev_trait,
                            is.na(trait) & !is.na(prev_trait) & !is.na(next_trait) & group == "A" ~ 
                                            next_trait,
                            is.na(trait) & !is.na(prev_trait) & !is.na(next_trait) & group == "B" ~ 
                                            prev_trait))

factor_output[,i] <- df$new_trait
}
```

# New data
```{r}
backfill_data <- traits %>% 
  select(anon_sname:statdate, na_count, yol, group) %>% 
  cbind(numeric_output) %>% 
  cbind(factor_output)
```


## Removing all rows (even finals) if there are more than 34 biomarkers missing
```{r}
backfill_data <- backfill_data %>% 
  filter(na_count < 34)
# removed 68 rows
# N = 2402

backfill_data %>% 
  ggplot() +
  geom_histogram(aes(x=na_count), binwidth = 1) +
  theme_bw()
```

## Removing unnecessary columns
```{r}
backfill_data <- backfill_data %>% 
  select(-c(birth:group))
```

